---
title: "fhimaps: map data for Norway"
output: rmarkdown::html_vignette
author: "Richard Aubrey White, Chi Zhang"
date: "2021-01-15"
vignette: >
  %\VignetteIndexEntry{fhimaps: map data for Norway}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```


`fhimaps` is a package for map visualisation developed by researchers at Norwegian Institute of Public Health (Folkehelseinstituttet, FHI). The package contains map data for Norway at different levels geographic granularity and layout. 


```{r echo = F, results='asis'}
tb <- readxl::read_xlsx(system.file("documentation", "fhimaps_ref.xlsx", package = "fhimaps"))
knitr::kable(tb)
```

## Naming scheme

The naming scheme has the following pattern: `scope_geogranularity_type_border_layout_Rclass`. 

#### Scope
This is what content the map data contains. Currently we focus on these 2 scopes: 

* country (Norway)
* city (Oslo)


#### Geogranularity 

This is the geographic granularity, which defines the border of county, municipality and ward. We follow the EU definition of territorial unit ([NUTS](https://ec.europa.eu/eurostat/web/nuts/nuts-maps), [LAU](https://ec.europa.eu/eurostat/web/nuts/local-administrative-units)) when applicable.

* [NUTS3](https://ec.europa.eu/eurostat/documents/345175/7451602/2021-NUTS-3-map-NO.pdf): Nomenclature of Territorial Unit level 3. Equivalent to fylke (county) in Norway.  

* LAU2: Local Administrative Unit level 2. Equivalent to kommune (municipality) in Norway. 

* ward: district within a municipality. Bydel in Norwegian. So far we only have the data for Oslo.

* xxx: only used for plotting `position_title`. (see example)


#### Type 

The type of the data object to distinguish whether it's a geographical map or coordinate for label positions. 

* map: map data, in either `data.table` or `sf` class.

* position geolabels: geographical coordinates for the position of labels, e.g. "county03" or "Oslo".

* position title (insert oslo): geographical coordinate for position of title. So far it's only for layout: `insert_oslo`. (see example)



#### Border

Due to recent redistricting, there exist multiple versions of county borders. We provide maps that match the borders in the following years: 

* 2020: the current border, this map contains 11 counties. 

* 2019: border before redistricting in 2020. This map contains 19 counties. 

* 2017: border before redistricting in 2018. 

More information on counties in Norway can be found [here](https://en.wikipedia.org/wiki/Counties_of_Norway#cite_note-13). 



#### Layout

Map layout. (easier to put a figure here)




#### R class

R class for the map object. 

* `data.table`: applicable for maps and label/title coordinates.

* `sf`: simple feature for spatial vector data. More on [sf](https://r-spatial.github.io/sf/articles/).



# Examples

```{r setup}
library(fhimaps)
library(ggplot2)
library(data.table)
library(magrittr)
```


## different layout

### default layout

```{r fig.height=6, fig.width=6}
pd <- fhimaps::norway_lau2_map_b2020_default_dt
q <- ggplot()
q <- q + geom_polygon(data = pd, aes( x = long, y = lat, group = group), 
                      color="black", fill="white", size=0.1)
q <- q + theme_void()
q <- q + coord_quickmap()
q <- q + labs(title = "")
q
```

### split 

```{r fig.height=6, fig.width=6}
pd <- fhimaps::norway_lau2_map_b2020_split_dt
q <- ggplot()
q <- q + geom_polygon(data = pd, aes( x = long, y = lat, group = group), 
                      color="black", fill="white", size=0.1)
q <- q + theme_void()
q <- q + coord_quickmap()
q <- q + labs(title = "")
q
```

### insert oslo

```{r fig.height=6, fig.width=6}
pd <- fhimaps::norway_lau2_map_b2020_insert_oslo_dt
q <- ggplot()
q <- q + geom_polygon(data = pd, aes( x = long, y = lat, group = group), 
                      color="black", fill="white", size=0.1)

# add title "Oslo"
q <- q + annotate(
  "text", 
  x = fhimaps::norway_xxx_position_title_insert_oslo_b2020_insert_oslo_dt$long, 
  y = fhimaps::norway_xxx_position_title_insert_oslo_b2020_insert_oslo_dt$lat,
  label = "Oslo"
  )
q <- q + theme_void()
q <- q + coord_quickmap()
q <- q + labs(title = "")
q
```

#### Oslo wards 

```{r fig.height=6, fig.width=6}
pd <- fhimaps::oslo_ward_map_b2020_default_dt
q <- ggplot()
q <- q + geom_polygon(data = pd, aes(x = long, y = lat, group = group), 
                      color="black", fill="white", size=0.1)
q <- q + theme_void()
q <- q + coord_quickmap()
q <- q + labs(title = "")
q
```


## customise the maps (color, label)

#### Automatic coloring by `location_code`

```{r fig.height=6, fig.width=6}
pd <- fhimaps::norway_nuts3_map_b2020_default_dt
q <- ggplot()
q <- q + geom_polygon(data = pd, aes(x = long, y = lat, group = group, fill = location_code),  
                      color="black", size=0.1)
q <- q + theme_void()
q <- q + coord_quickmap()
q <- q + labs(title = "")
q
```
 
#### Customized coloring 


```{r fig.height=6, fig.width=6}
pd <- fhimaps::norway_nuts3_map_b2020_default_dt

# assign each location a random category for different colors
location_info <- unique(pd[,c("location_code")])
suppressWarnings(
  location_info[,category:=rep(c("Good","Normal","Neutral","Bad","Very Bad"),
                               each=3)[1:.N]])
location_info[,category:=factor(
  category,levels=c("Good","Normal","Neutral","Bad","Very Bad"))]
print(location_info)

# join the map data.table
pd[location_info,on="location_code",category:=category]

q <- ggplot()
q <- q + geom_polygon(data = pd, aes( x = long, y = lat, group = group, fill=category), 
                      color="black", size=0.25)
q <- q + coord_quickmap()
# it is possible to use the color palette provided in fhiplot
# q <- q + fhiplot::scale_fill_fhi("Category",palette = "map_seq_complete", direction = 1)
q <- q + labs(title="")
q <- q + theme_void()
q
```

# 

```{r fig.height=6, fig.width=6}
pd <- fhimaps::norway_nuts3_map_b2020_default_dt
q <- ggplot()
q <- q + geom_polygon(data = pd, aes(x = long, y = lat, group = group, fill = location_code),  
                      color="black", size=0.1)
q <- q + geom_label(
  data = fhimaps::norway_nuts3_positions_geolabels_b2020_default_dt,
  mapping = aes(x = long, y = lat, label = location_code)
  )
q <- q + theme_void()
q <- q + coord_quickmap()
q <- q + labs(title = "")
q

# ggrepel::geom_label_repel() for avoiding overlap
```


```{r fig.height=6, fig.width=6}
q <- ggplot(mapping = aes(x = long, y = lat))
q <- q + geom_polygon(
   data = fhimaps::oslo_ward_map_b2020_default_dt,
   mapping = aes(group = group),
   color = "black",
   fill = "white",
   size = 0.2
 )
 q <- q + ggrepel::geom_label_repel(
   data = fhimaps::oslo_ward_positions_geolabels_b2020_default_dt,
   mapping = aes(label = location_code),
   color = "red"
 )
 q <- q + theme_void()
 q <- q + coord_quickmap()
 q

```

### enrich plot with additional data 



```{r fig.height=6, fig.width=6}
label_county <- fhimaps::norway_nuts3_positions_geolabels_b2020_default_dt

# enrich with population and location name
dpop_county_2020 <- fhidata::norway_population_b2020[
  granularity_geo == 'county' & year == '2020'][
    , pop_total := sum(pop), by = location_code][
      , .(location_code, pop_total)] %>% unique

dloc_county_2020 <- fhidata::norway_locations_long_b2020[granularity_geo == 'county',
                                             .(location_code, location_name)]

# join, create label
dd <- label_county[dpop_county_2020, on = 'location_code']
dd <- dd[dloc_county_2020, on = 'location_code']
dd[, label := paste0(location_name, '\n', pop_total)]
print(head(dd))

# plot
pd <- fhimaps::norway_nuts3_map_b2020_default_dt
q <- ggplot()
q <- q + geom_polygon(data = pd, aes(x = long, y = lat, group = group, fill = location_code),
                      color="black", size=0.1)
q <- q + ggrepel::geom_label_repel(
  data = dd,
  mapping = aes(x = long, y = lat, label = label)
)
q <- q + theme_void()
q <- q + coord_quickmap()
q <- q + labs(title = "")
q

```



```{r fig.height=6, fig.width=6}
label_oslo <- fhimaps::oslo_ward_positions_geolabels_b2020_default_dt

# enrich with population and location name
dpop_oslo_2020 <- fhidata::norway_population_b2020[
  granularity_geo == 'wardoslo' & year == '2020'][
    , pop_total := sum(pop), by = location_code][
      , .(location_code, pop_total)] %>% unique


dloc_oslo_2020 <- fhidata::norway_locations_long_b2020[granularity_geo == 'wardoslo',
                                             .(location_code, location_name)]

# join, create label
dd <- label_oslo[dpop_oslo_2020, on = 'location_code']
dd <- dd[dloc_oslo_2020, on = 'location_code']
dd[, label := paste0(location_name, '\n', pop_total)]
print(head(dd))

q <- ggplot(mapping = aes(x = long, y = lat))
q <- q + geom_polygon(
  data = fhimaps::oslo_ward_map_b2020_default_dt,
  mapping = aes(group = group),
  color = "black",
  fill = "white",
  size = 0.2
)
q <- q + ggrepel::geom_label_repel(
  data = dd,
  mapping = aes(label = label),
  color = "red"
)
q <- q + theme_void()
q <- q + coord_quickmap()
q

```


## `sf` for interactive 


```{r fig.height=6, fig.width=6}
pd_county <- fhimaps::norway_nuts3_map_b2019_default_sf
leaflet::leaflet(
  pd_county,
  options = leaflet::leafletOptions(preferCanvas = F)
) %>% 
  leaflet::addProviderTiles(leaflet::providers$Esri.WorldGrayCanvas) %>%
  leaflet::addPolygons(
   #  fillColor = ~ pal_not_censored(n_0_13_fixed_status_with_numbers),
    weight = 0.3,
    opacity = 1,
    color = "black", fillColor = 'white',
    fillOpacity = 0.9  )
```



# to do 
2. `norway_nuts3_positions_geolabels_b2020_insert_oslo_dt`, change oslo coordinate



